### 2017-01-31 local-capability-machine.tex
typo
Some work on FTLR proof.  Move some more lemmas forward.

%% make latex preview-mode work with natbib...
%\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}

@@

@@

\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
  \label{lem:writecond-mono-priv}
  If
  \begin{itemize}
  \item $W' \futurestr W$
  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
  \end{itemize}
  then
  \[
    \npair{\(start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
  % TODO
\end{proof}

\subsubsection{LR Sanity lemmas}

\begin{lemma}[Heap satisfaction downwards closure]
  \label{heap-sat-dc}
  \begin{align*}
    &\forall \hs, n' \leq n, W \ldotp \\
    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
  \end{align*}
\end{lemma}

@@

\begin{lemma}[Fundamental theorem proof helper lemma]
  \label{lem:ftlr-induction-helper}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
  If one of the following holds:
  \begin{itemize}
  \item \[
      \begin{gathered}
        \perm = \exec \land\\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{gathered}
    \]
  \item \[
      \begin{gathered}
        \perm = \rwx \land \\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{gathered}
    \]
  \item \[
      \begin{gathered}
        \perm = \rwlx \land\\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{gathered}
    \]
  \end{itemize}
  and
  \begin{itemize}
  \item $\npair[n]{\reg}\in \stdrr(W)$
  \item $\heapSat[\hs]{n}{W}$
  \item $i \leq n$
  \item $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$
  \end{itemize}

  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
  $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
\end{lemma}
\begin{proof}
  \begin{enumproof}
  \item By induction on $n$. In other words, assume that the theorem already
    holds for all $n' < n$.
  \item $i \neq 0$, since
    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
    Therefore, assume w.l.o.g. that $i = 1+i'$,
    \begin{equation*}
      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
    \end{equation*}
  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
    and $\var{conf'} =
    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
    \uplus \ms_f))$\\
    By definition of
    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
    is not possible that $\failed \step[i'] (\halted,\Phi)$)
  \item We do a case analysis on
    $\memheap(a)$
    \begin{enumproof}
    \item Case $\memheap(a) = \jmp{\lv}$:
    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
    \item Case $\memheap(a) = \move{\lv}{\rv}$:
    \item Case $\memheap(a) = \load{\lv}{\hv}$:
    \item Case $\memheap(a) = \store{\hv}{\rv}$:
    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
    \item Case $\memheap(a) = \fail$: Contradiction, since then
      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
      (\halted,\Phi)$ which is not possible.
    \item Case $\memheap(a) = \halt$:
      \begin{enumproof}
      \item
        $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
        \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
        \uplus \ms_f))$\\
        By definition of $\sem{\cdot}$
      \item Then $i' = 0$ and $\Phi =
        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
        \ms_f)$. Follows from $(\halted,
        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
        \ms_f)) \step[i'] (\halted,\Phi)$
      \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
        $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
        $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
      \end{enumproof}
    \end{enumproof}
  \end{enumproof}
\end{proof}

>>>>

%make latex preview-mode work with natbib...
\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}

@@

\subsubsection{Downward closedness lemmas}
\label{sec:downward-closedness}

\begin{lemma}[$\stdvr{\cdot}$ downwards-closed]
  \label{lem:stdvr-dc}
  \begin{itemize}
  \item $n' \leq n$
  \item $\npair{\word} \in \stdvr(W)$
  \end{itemize}
  then 
  \[
   \npair[n']{\word} \in \stdvr(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:stdvr-dc}]
  % TODO
\end{proof}

\begin{lemma}[$\stdrr{\cdot}$ downwards-closed]
  \label{lem:stdrr-dc}
  \begin{itemize}
  \item $n' \leq n$
  \item $\npair{\reg} \in \stdrr(W)$
  \end{itemize}
  then 
  \[
   \npair[n']{\reg} \in \stdrr(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:stdrr-dc}]
  % TODO
\end{proof}

\begin{lemma}[Heap satisfaction downwards closure]
  \label{lem:heap-sat-dc}
  \begin{align*}
    &\forall \hs, n' \leq n, W \ldotp \\
    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
  \end{align*}
\end{lemma}

\subsubsection{Monotonicity lemmas}
\label{sec:mono}

\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
  \label{lem:writecond-mono-priv}
  If
  \begin{itemize}
  \item $W' \futurestr W$
  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
  \end{itemize}
  then
  \[
    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
  % TODO
\end{proof}

@@

@@

\begin{lemma}[Conditions for load instruction are sufficient]
  \label{lem:conds-load-suffice}
  If
  \begin{itemize}
  \item $\heapSat[\memheap]{n}{W}$
  \item $\var{c} = ((\perm',\gl'),\start',\addrend',\addr')$
  \item $\npair{c}\in\stdvr(W)$
  \item $\readAllowed{\perm'}$ 
  \item $\withinBounds{\var{c}}$
  \end{itemize}

  then $\npair{\memheap(\addr')} \in \stdvr(W)$
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}[Conditions for lea instruction are sufficient]
  \label{lem:conds-lea-suffice}
  If
  \begin{itemize}
  \item $\npair{((\perm,\gl),\start,\addrend,\addr)}\in\stdvr(W)$
  \item $\perm \neq \entry$
  \end{itemize}

  then $\npair{((\perm,\gl),\start,\addrend,\addr')} \in \stdvr(W)$
\end{lemma}
\begin{proof}
  Follows by inspection of the cases in the definition of $\stdvr(W)$: $a$ is
  ignored in all cases except where $\perm = \entry$.
\end{proof}

\begin{lemma}[Conditions for restrict instruction are sufficient]
  \label{lem:conds-restrict-suffice}
  If
  \begin{itemize}
  \item $\npair{((\perm,\gl),\start,\addrend,\addr)}\in\stdvr(W)$
  \item $(\perm',\gl')\sqsubseteq (\perm,\gl)$
  \end{itemize}

  then  $\npair{((\perm',\gl'),\start,\addrend,\addr)}\in \stdvr(W)$
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}[Conditions for subseg instruction are sufficient]
  \label{lem:conds-subseg-suffice}
  If
  \begin{itemize}
  \item $\npair{((\perm',\gl'),\start',\addrend',\addr')} \in \stdvr(W)$
  \item $\start' \leq \start''$
  \item $\addrend'' \leq \addrend'$
  \end{itemize}

  then $\npair{((\perm',\gl'),\start'',\addrend'',\addr')} \in \stdvr(W)$
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}[Conditions for store instruction are sufficient]
  \label{lem:conds-store-suff}
  If 
  \begin{itemize}
  \item $\heapSat[m]{n}{W}$
  \item $((\perm,\gl),\start,\addrend,\addr) = c$
  \item $\npair{c}\in\stdvr(W)$
  \item $\writeAllowed{\perm}$
  \item $\withinBounds{\var{c}}$
  \item $\npair{\var{w}}\in\stdvr(W)$
  \item if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in
    \{\rwlx,\readwritel \}$
  \end{itemize}

  then $\heapSat[{m\update{\addr}{\var{w}}}]{n}{W}$
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}[Fundamental theorem proof helper lemma]
  \label{lem:ftlr-induction-helper}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
  If one of the following holds:
  \begin{itemize}
  \item \[
      \begin{gathered}
        \perm = \exec \land\\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{gathered}
    \]
  \item \[
      \begin{gathered}
        \perm = \rwx \land \\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{gathered}
    \]
  \item \[
      \begin{gathered}
        \perm = \rwlx \land\\
        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{gathered}
    \]
  \end{itemize}
  and
  \begin{enumproof}
  \item $\npair[n]{\reg}\in \stdrr(W)$
  \item $\heapSat[\hs]{n}{W}$
  \item $i \leq n$
  \item $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)$
  \item $\Phi \step[i] (\halted,\Phi')$
  \end{enumproof}

  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
  $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
\begin{proof}
  \item By induction on $n$. In other words, assume that the theorem already
    holds for all $n' < n$.
  \item $i \neq 0$, since
    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
    \ms_f)\neq (\halted,\Phi')$ for any $\Phi'$. \\
    Therefore, assume w.l.o.g. that $i = 1+i'$,
      \Phi \step \var{conf'} \step[i'] (\halted,\Phi')
  \item $n > 0$, since otherwise $i = 0$ (because $i \leq n$) and this is
    impossible by the previous point.
  \item By inspection of the definitions of $\Phi\step \var{conf'}$ and
    $\sem{\decode(\memheap(\addr))}$ and $\updatePcPerm{\cdot}$ and
    $\stdUpdatePc{\cdot}$, it is easy to see that one of the following cases
    must hold:
  \item Case $\var{conf'} = \failed$: contradiction, since it is not possible
    that $\failed \step[i'] (\halted,\Phi)$. 
  \item Case $\var{conf'} = (\halted,\Phi)$:
    \item Then $i' = 0$ and $\Phi' =
      \Phi$\\
      Follows from $(\halted,
      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
      \ms_f)) \step[i'] (\halted,\Phi)$
    \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
      $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and
      $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
  \item Case $\var{conf'} = \Phi'$, and additionally, one of the following
    holds:
    \item $\memheap[\Phi'] = \memheap[\Phi]$
    \item $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with
      $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$ and
      $\writeAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
      \memreg(r_2)$ and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm' \in
      \{\rwlx,\readwritel \}$
    and also one of the following holds:
    \item $\memreg[\Phi'](\pcreg) = \updatePcPerm{\memreg(\lv)}$
    \item $\memreg[\Phi'](\pcreg) = \var{newPc}$ and $\var{newPc} =
      ((\perm,\gl),\start,\addrend,\addr + 1)$
      \dominique{This is wrong: the semantics allow, for example, a move into
        the PC, after which the new pc is increased by one}
    and finally, for all $r \in (\RegName \setminus \{\pcreg\})$, one of the following
    holds:
    \item $\memreg[\Phi'](r) = \memreg[\Phi](r)$
    \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$
    \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
      ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
      $\readAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
      \memheap(\addr')$
    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
      ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
      \ints$
    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
      ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
      (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$
    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
      ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
      $\addrend'' \leq \addrend'$ and $c =
      ((\perm',\gl'),\start'',\addrend'',\addr')$
    In this case, we have:
    \item $\heapSat[{\memheap[\Phi']}]{n-1}{W}$.
      \item Case $\memheap[\Phi'] = \memheap[\Phi]$:
        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
      \item Case $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$,
        with $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$
        and $\writeAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
        \memreg(r_2)$ and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm'
        \in \{\rwlx,\readwritel \}$.\\
        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by
        Lemmas~\ref{lem:conds-store-suff} and~\ref{lem:heap-sat-dc} using the
        fact that $\heapSat[\memheap]{n}{W}$ and
        $\npair{\memreg(r_1)}\in\stdvr(W)$ and
        $\npair{\memreg(r_2)}\in\stdvr(W)$ which follows from
        $\npair{\memreg}\in\stdrr(W)$ by definition.
    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$.
      \item Assume: $r \in \RegName\setminus\{\pcreg\}$.\\
        Suffices: $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$.\\
        By: definition of $\stdrr{\cdot}$
      \item Case $\memreg[\Phi'](r) = \memreg[\Phi](r)$:
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows from
        $\npair{\memreg}\in\stdrr(W)$ by definition of $\stdrr(\cdot)$ using
        Lemma~\ref{lem:stdvr-dc}.
      \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$.
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by definition of
        $\stdvr(\cdot)$
      \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
        ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
        $\readAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
        \memheap(\addr')$:\\
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
        Lemmas~\ref{lem:conds-load-suffice} and~\ref{lem:stdvr-dc} using the
        fact that $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
        ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
        $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
        \ints$: \\
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
        Lemmas~\ref{lem:conds-lea-suffice} and~\ref{lem:stdvr-dc} using the
        fact that $\npair{\memreg(r_1)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
        ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
        (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$: \\
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
        Lemmas~\ref{lem:conds-restrict-suffice} and~\ref{lem:stdvr-dc} using the
        fact that $\npair{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by
        definition\dominique{this does not follow if $r = \pcreg$, similar
          problem in other cases}.
      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
        ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
        $\addrend'' \leq \addrend'$ and $c =
        ((\perm',\gl'),\start'',\addrend'',\addr')$: \\
        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
        Lemmas~\ref{lem:conds-subseg-suffice} and~\ref{lem:stdvr-dc} using the
        fact that $\npair{\memreg(r)}\in\stdvr(W)$
        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
    \item $\npair[n-1]{\memreg[\Phi'](\pcreg)} \in \stder(W)$: TODO
